### 2018-10-30 paper/reply.tex
PAPER: begin changes requested by referee

\documentclass[fontsize=12pt, paper=a4]{scrlttr2}

% Dont forget to read the KOMA-Script documentation, scrguien.pdf

\setkomavar{fromname}{Andrew Nelson} % your name
\setkomavar{fromaddress}{ANSTO \\ Locked Bag 2001 \\ Kirrawee DC \\ NSW 2232 \\ Australia}

\setkomavar{signature}{} % printed after the \closing
\renewcommand{\raggedsignature}{\raggedright} % make the signature ragged right

\setkomavar{subject}{RG5158 - refnx -- Neutron and X-ray reflectometry analysis in Python} % subject of the letter

\begin{document}
\begin{letter}{Dr Gilles Renaud \\ gilles.renaud@cea.fr}

\opening{Dear Dr Renaud}  % eg. Hello
Thank you for forwarding the kind comments of the referees towards my recently submitted manuscript (\#RG5158). I note that they were universally positive, with only minor adjustments to be made. These adjustments towards each referees comments are listed below, and I trust that they meet the requirements for publication in the \emph{Journal of Applied Crystallography}.

\textbf{referee 1}

\textbf{referee 2}
\begin{itemize}
\item \textbf{Resolve the question marks in the Sivia and Skilling reference} These references are completed successfully in our rendering of the latex document. I am unsure why they are not rendered properly by the submission gateway.
\item \textbf{In middle of page, Replace "user user" with "user"} This change is made.
\item \textbf{Third-to-last line – shouldn't this read as p(&thetav;|D,I) instead of p(&thetav;|I,D)?} The order of D, I does not really matter, but we have changed it anyway to be consistent with equation 1.
\end{itemize}

\textbf{referee 3}
No amendments are required.

\textbf{referee 4}
\begin{itemize}
\item \textbf{Missing references in 1-Introduction (after Sivia 2006)?} These references are completed successfully in our rendering of the latex document. I am unsure why they are not rendered properly by the submission gateway.
\item \textbf{Pauw et al may be added to the references list} This reference has been added.
\item \textbf{please complete the reference Nelson, A and Dewhurst C 2014 46 1338-1346} This reference has been completed.
\end{itemize}

\closing{Regards, \\ Dr Andrew Nelson} %eg. Regards

\end{letter}
\end{document}